\chapter{Acknowledgements}

\footnotesize
As it turns out, completing a dissertation is \textit{a lot} of work. I am eternally grateful to the large number of people who have helped me along the way, and without whom this work would not have been possible. I can assure you if I were to properly thank everyone who has helped me, this acknowledgements section would be longer than the ``systematic studies'' chapter of this dissertation. So, I will try to keep it brief, but even by virtue of you reading this sentence, you have my sincerest thanks.

To my advisor, Professor Christina Markert, for her guidance and support throughout my graduate career. I am grateful for the opportunity to have worked with her, and for all of the time and effort she has put into helping me succeed, both as a graduate student \textit{and} as a scientist. To Amanda and Josey, who have had to share an office with me for most of their time at UT. I appreciate all of the conversations and friendship we've shared over the years. To Erin and Alex, who welcomed me into the group and immediately made me feel at home. To Professor Deepa Thomas, who works harder than anyone I have ever met, and yet \textit{somehow} still managed to find time to answer my endless questions. To Lanny, who always made me \textit{believe} I'm capable of understanding the theory of our field. I'm not, but he still made me feel like I could. To Justin, who I've spent more time bugging in his office about random puzzles than I've spent writing this document. His seemingly infinite patience and willingness to help me with my many, many struggles has been invaluable.

To all of the friends I made within the UT Counter Strike community. Without them, I would have finished this dissertation \textit{much} sooner, but all of the memories we've made over the years have been well worth it. To Devanshu and Richard, who would brighten up my day with horrible puns and surprise visits to my office. To my roommates Travis, Kevin and Luke, whose friendship has pulled me through some of the most stressful times. To Matteo, Arild, Matthias and Jo, who converted me from someone who had never heard the acronym ``ssh'' into a full-fledged computer nerd. Without their endless support during my times at CERN, I would have been completely lost. 

To my family, especially my parents and sister, whose love and support during this entire process (a.k.a. my life) cannot be overstated. Their encouragement and belief in me as I pursue my passions has been a constant source of strength. To my fiance\'e Jaynee, who has given me enough happiness and love to last a lifetime, which is \textit{just} enough to offset the stress and anxiety of writing a dissertation. I am so very much excited to spend the rest of my life with her. To Jaynee's family, who have welcomed me into their family with open arms and given me plenty of indispensable guidance during these past few years.

\normalsize